Supermartingales and Termination Analysis of Statistical PCF

### 2020-08-30 probabilisticProgrammingMartingales.tex
Terms -> skeletons: some changes and an example. Def. of I^{L_s(M)}

\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{enumitem}
\newcommand{\tY}{\textsf{Y}}
%\newcommand{\tif}[3]{\textsf{if }#1\textsf{ then }#2\textsf{ else }#3}
\newcommand{\tif}[3]{\textsf{if}\,(#1, #2, #3)} % space-saving concrete syntax
\newcommand{\tsample}{\textsf{sample}}
\newcommand{\tscore}{\textsf{score}}
\newcommand{\skeletonPlaceholder}{\textsf{X}} % I'm not sure what the best notation here is.
\DeclareMathOperator{\red}{red}
\DeclareMathOperator{\nnext}{next}
\DeclareMathOperator{\cbv}{cbv}

@@

\subsection{Sampling semantics}
This version of the reduction relation allows $\tsample$ to reduce to any number in $[0,1]$. To more precisely specify the probabilities, an additional argument is needed to determine the outcome of random samples. Let $ I = [0,1] \subset \mathbb{R} $, and let $S = I^{\mathbb{N}}$, with the \changed[lo]{Borel $\sigma$-algebra $\calF$} and the probability measure, denoted $\mu$, given by the limit of $1 \gets I \gets I^2 \gets \cdots$, where the maps are the projections that ignore the last element. Equivalently, a basis of measurable sets is $\prod_{i=0}^\infty X_i$ where $X_i$ are all Borel and all but finitely many are $I$, and $\mu (\prod_{i=0}^\infty X_i) = \prod_{i=0}^\infty \Leb(X_i)$.
The maps $\pi_h:S \to I, \; \pi_t:S \to S$ popping the first element are then measurable.
\changed[lo]{Following \cite{DBLP:conf/esop/CulpepperC17}, we call the probability space $(S, \calF, \mu)$ the \emph{entropy space}.}

@@

A \emph{position} 
\akr{I'm not really sure how to phrase the citation, but this is adapted from \cite{Kennaway96infinitarylambda}.} 
\lo{The idea is common enough. I think no citation necessary.}
is a finite sequence of steps into a term, defined inductively as
\begin{align*}
P ::= \cdot \mid \lambda ; P \mid @_1 ; P \mid @_2 ; P \mid \underline f_i ; P \mid \tY ; P \mid \textsf{if}_1 ; P \mid \textsf{if}_2 ; P \mid \textsf{if}_3 ; P \mid \tscore ; P.
\end{align*}
The \emph{subterm of $M$ at a position $P$}, denoted $M \mid P$, is defined as
\begin{align*}
M \mid \cdot & = M \\
\lambda x. M \mid \lambda ; P & = M \mid P \\
M_1 M_2 \mid @_i ; P & = M_i \mid P \quad \text{for } i = 1,2 \\
\underline f(M_1,\dots,M_n) \mid \underline f_i ; P & = M_i \mid P \quad \text{for }i \leq n \\
\tY M \mid \tY ; P & = M \mid P \\
\tif{M_1 < 0}{M_2}{M_3} \mid \textsf{if}_i ; P & = M_i \mid P \quad \text{for } i = 1,2,3 \\
\tscore(M) \mid \tscore ; P & = M \mid P
\end{align*}
so that every subterm is located at a unique position, but not every position corresponds to a subterm (e.g. $x \, y \mid \lambda ; \cdot$ is undefined). 
A position such that $M\mid P$ does exist is said to occur in $M$. 
\changed[lo]{\emph{Substitution} of $N$ at position $P$ in $M$, written $M[N/P]$, is defined similarly.
For example 
\begin{align*}
M &= \lambda x \, y. y \, (\tif{x < 0}{y \, (\underline f (x))}{\underline 3})\\
P &=\lambda ; \lambda ; @_2 ; \textsf{if}_2 ; @_2
\end{align*} 
then 
\(
M[\tsample / P] = \lambda x \, y. y \, (\tif{x < 0}{y \, \tsample}{\underline 3}).
\)}

Two subterms $N_1$ and $N_2$ of a term $M$, corresponding to positions $P_1$ and $P_2$, can overlap in a few different ways. 
If $P_1$ is an initial segment (i.e.~prefix) of $P_2$ (written as $P_1 \leq P_2$), then $N_2$ is also a subterm of $N_1$. If neither $P_1 \leq P_2$ nor $P_1 \geq P_2$, the positions are said to be disjoint. 
This is mostly relevant in that in any substitution at some position, the subterms at all other positions are unaffected. \lo{The preceding sentence is unclear. Typo? Mostly $\mapsto$ most?}

With this notation, a more general reduction relation $\to$ can be defined. \akr{It would almost work to have one of the neater versions of $\tY$-reduction here.}
%\lo{On first reading, I struggled to parse the following clauses.}
\begin{definition}
\label{def:more general red}
\changed[lo]{The binary relation $\to$ is defined by the following rules, each is conditional on a redex occurring at position $P$ in the term $M$.}
\begin{align*}
  \text{if } M \mid P = (\lambda x.s) v,\ & M \to M[s[v/x]/P] \\
  \text{if } M \mid P = \underline f (\underline r_1, \dots , \underline r_n),\ & M \to M[\underline{f(r_1,\dots,r_n)}/P] \\
  \text{if } M \mid P = \tY \lambda x. s,\ & M \to M[\lambda z. s[(\tY \lambda x. s)/x] z/P] \text{ where $z$ is not free in $s$}\\
  \text{if } M \mid P = \tif{\underline r < 0}{s_1}{s_2},\ & M \to M[s_1/P] \text{ where }r < 0 \\
  \text{if } M \mid P = \tif{\underline r < 0}{s_1}{s_2},\ & M \to M[s_2/P] \text{ where }r \geq 0 \\
  \text{if } M \mid P = \tsample & \text{ and $\lambda$ does not occur after $@_2$ or $\tY$ in $P$},\\ & M \to M[\underline r/P] \text{ where } r \in [0,1] \\
  \text{if } M \mid P = \tscore(\underline r),\ & M \to M[\underline r/P].
\end{align*}
In each of these cases, $M \mid P$ is the redex, and reduction takes place at $P$.
\end{definition}

\paragraph{}
Labelling the pre-chosen samples by the positions in the term would also not work because in some cases, a $\tsample$ will be duplicated before being reduced, for example, in $(\lambda x. x \, {\underline 0} \, \mathbin{\underline{+}} \, {x \, \underline 0}) (\lambda y. \tsample)$, both of the $\tsample$ redexes that eventually occur originate at $@_2 ; \lambda$. It is therefore necessary to consider possible positions that may occur in other terms reachable from the original term. This is itself inadequate because some of the positions in different reachable terms need to be considered the same, and the number of reachable terms is in general uncountable, which leads to measure-theoretic issues.

Define a \emph{skeleton} to be a term but, instead of having real constants $\underline r$, it has a placeholder $\skeletonPlaceholder$, so that each term has a skeleton and each skeleton can be converted to a term given the appropriate number of real numbers to substitute in. The positions in a skeleton and the reduction relation on skeletons can be extended from the definitions on terms in the obvious way, with $\tif{\skeletonPlaceholder}{A}{B}$ reducing nondeterministically to both $A$ and $B$.

Given a closed term $M$, let $L_0(M)$ be the set of pairs, the first element of which is a reduction sequence of skeletons starting at $M$, and the second of which is a position in the final skeleton of the reduction sequence. \akr{I've yet to go through and replace references to terms with references to skeletons, but it should be reasonably obvious where the substitution should take place. This is potentially even more confusing though because I'm now implicitly identifying terms, skeletons, reduction sequences and skeletal reduction sequences.} Reduction sequences are used rather than reachable terms because if the same term is reached twice, different samples may be needed, but they will be discussed as though they were terms, by abuse of notation. \lo{Readers would benefit greatly from some examples of an $L_0(M)$-pair, and the relations $\sim_p$ and $\sim_c$, here, before the formal definitions. Explain the underlying intuition.} The relation $\sim$ is defined as the union of the symmetric relations $\sim_p$ \changed[lo]{($p$ for parent-child relationship)} and $\sim_c$ \changed[lo]{($c$ for cousin relationship)} where
\lo{As defined, $\sim_p$ is not a symmetric relation.}
\begin{enumerate}
    \item If $N$ reduces to $O$ with the redex at position $P$ \changed[lo]{(i.e.~$M \to^\ast N \to O = N[\dotDelta / P]$ with $O \mid P = \Delta$ for some redex contraction $\Delta \to \dotDelta$)}, and $Q$ is a position in $N$ disjoint from $P$, then $(N,Q) \sim_p (O,Q)$.

@@

        \item $N \mid P = \tY (\lambda x. A)$, $A$ reduced at $Q$ is $A'$, $(R_i)_i$ is a list of all the positions where $A' \mid R  = x$, ordered from left to right, $O_1$ is $N$ reduced first at $P;\tY;\lambda;Q$ then at $P$, and $O_2$ is $N$ reduced first at $P$ then at $P;\lambda;@_1;Q$ then at $P;\lambda;@_1;R_i;Y;\lambda;Q$ for each $i$ in order.
    \end{enumerate}
    (in which case $O_1$ and $O_2$ are equal as terms, but with different reduction sequences), $O_1'$ and $O_2'$ are the results of applying some reduction sequence to each of $O_1$ and $O_2$ (the same reductions in each case, which is always possible because they're equal terms), and $S$ is a position in $O_1'$ (or equivalently $O_2'$), then $(O_1',S) \sim_c (O_2',S)$.
\end{enumerate}

@@

\begin{example}
See \Cref{fig:adk-diagram} for a graphical depiction of $\sim_p$- and $\sim_c$-rules.
\end{example}

The reflexive transitive closure $\sim^*$ of this relation is used to define the set of \emph{potential positions} $L(M) = L_0(M) / \sim^*$, and each equivalence class can be considered as the same position as it may occur across multiple reachable terms. 
If $(N,P) \sim^* (O,Q)$, then $N \mid P$ and $O \mid Q$ both have the same shape (i.e.~they're either both variables, both applications, both $\tsample$s etc.), therefore it's well-defined to talk of the set of potential positions where there is a $\tsample$, $L_s(M)$. 
\lo{Thus $L_s(M) := \{[(X, P)]_{\sim^\ast_M} : X \mid P = \tsample\}$.}
The new sample space is then defined as $I^{L_s(M)}$, with the Borel $\sigma$-algebra and product measure.

\paragraph{}
For now, here's a summary of the lemmas and theorems I intend to prove about this alternative semantics:
\begin{itemize}
    \item If $M \to N$, the relation $\sim_N^*$ defined on $Rch(N)$ is equal to the restriction of $\sim_M^*$ to $Rch(N)$ %\lo{Define $\sim_N^*$.}
    \item Given the definition below of $\Rightarrow$, which is still applicable to this version of $\sim$, and which is the equivalent of $\red$ in this version of the semantics, $\Rightarrow$ is Church-Rosser.
    \item If $M \to N$ at $P$, no position in any term in $Rch(N)$ is related to $(M,P)$ by $\sim_M^*$, and therefore each sample is used at most once in any reduction sequence.
    \item The subprobability distribution of values obtained by reducing a term in the usual call-by-value semantics is equal to that obtained by any reduction strategy on $\Rightarrow$ (i.e.~a sub-relation of $\Rightarrow$ that is a function) (with some allowance made for the fact that some values are not normal forms of $\Rightarrow$, so the reduction strategy might overshoot the value that $\red^\infty$ reaches. I'm not quite sure what form this will take.)
    \item Similarly, the probability of termination is equal in $\red^\infty$ and any reduction strategy on $\Rightarrow$, so almost sure termination can be equivalently defined based on $\Rightarrow$.
    \item For any closed term $N$, if there is any reduction strategy for $N$ and ranking function corresponding to that reduction strategy, $N$ is AST.
    \item Corresponding sparse and antitone-strict ranking function theorems.
\end{itemize}

@@

\begin{lemma}
The relation $\sim$ is defined on $L_0(M)$ with reference to a particular starting term $M$, so different versions, $\sim_M$ and $\sim_N$, can be defined starting at different terms. If $M \to N$, then $\sim^*_N$ is equal to the restriction of $\sim^*_M$ to $L_0(N)$.
\end{lemma} 
\begin{proof}
$\sim_N^*$ is trivially a subset of $\sim_M^*$ because $\sim_N$ is a subset of $\sim_M$.

@@

\begin{definition}
A position in some (reduction sequence of) term(s) is \emph{derived from} another if it is related by the reflexive transitive closure of $\rightsquigarrow$, where $(V,R) \rightsquigarrow (W,T)$ just if $V \to W$ and one of the following cases holds:
\begin{itemize}
    \item $(V,R) \sim_p (W,T)$
    \item $R = T$, $V \to W$ at $A$, and $A > R$
    \item $V | A = (\lambda x.U) Z, U | B = x, R = A;@_2;C$ and $T = A;B;C$
    \item $V | A = \tY (\lambda x.U), R = A;\tY;\lambda;B$ and $T = A;\lambda;@_1;B$
    \item $V | A = \tY (\lambda x.U), U | B = x, R = A;C$ and $T = A;\lambda;@_1;B;C$
\end{itemize}
\end{definition}

@@

At each reduction step $M \to N$, the sample space must be restricted from $I^{L_s(M)}$ to $I^{L_s(N)}$. The injection $L_0(N) \to L_0(M)$ is trivial to define by appending $M \to N$ to each path, and using the previous lemma, this induces a corresponding injection on the quotient, $L(N) \to L(M)$. The corresponding map $L_s(N) \to L_s(M)$ is then denoted $i(M \to N)$.

@@

Let $I$ be $N$ reduced at $P$, and $J$ be $N$ reduced at $P;@_1;\lambda;Q$, so that $N \to I \to O_1$ and $N \to J \to O_2$. All of the reduction positions are $\geq P$, and $G$ is not disjoint from all of them, therefore $G$ is not disjoint from $P$. Let $K$ be the position such that $(N,G) \sim_p (I,K) \sim_p (O_1,H)$. The fact that $(N,G) \sim_p (I,K)$ implies that $G > P;@_1;\lambda$. Let $G = P;@_1;\lambda;L$ and $H = P;L'$. If $L$ is disjoint from $Q$, then $(N,G) \sim_p (J,G) \sim_p (O_2,P;L') = (O_2;H)$ by cases 1 and 2 of $\sim_p$. In the other case, that $L$ isn't disjoint from $Q$, $L > Q$ because none of the positions $<= P;@_1;\lambda;Q$ in $N$ are related to any position in $I$ by $\sim_p$. As $(I,P;L) \sim_p (O_1,P;L')$ (with the redex at $P;Q$), for exactly the same reason $(N,P;@_1;\lambda;L) \sim_p (J,P;@_1;\lambda;L')$ (with the redex at $P;@_1;\lambda;Q$). Because $(N,P;@_1;\lambda;L) \sim_p (J,P;@_1;\lambda;L')$, $N|P;@_1;\lambda;L = J|P;@_1;\lambda;L'$, and $N|P;@_1;\lambda;L \neq $ the variable of $N|P;@_1$, therefore $J|P;@_1;\lambda;L'$ is also not the variable therefore $(J,P;@_1;\lambda;L') \sim_p (O_2, P;L') = (O_2, H)$ by case 2 of $\sim_p$. Combining these results, $(N,G) \sim_p (O_2;H)$ as desired.

\begin{lemma} \label{redexDestroyed}
If $M \to N$, with the redex at position $P$, then no position in any term reachable from $N$ is related by $\sim^*$ to $(M,P)$.
\end{lemma}
\begin{proof}
Suppose on the contrary that $(M,P) \sim^* (A,D)$, where $N \to^* A$, then by Lemma \ref{simpDoesn'tMix}, $(M,P) \sim_p^* (A,D)$. $M \neq A$ therefore $(M,P) \sim_p$ some position in $N$, but in all the cases of the definition of $\sim_p$, no position in the child term is related to the position of the redex.
\end{proof}

\paragraph{}
The reduction relation $\Rightarrow$ is nondeterministic, so it admits multiple possible reduction strategies. A reduction strategy starting from a closed term $M$ is a partial function $f$ from $Rch(M)$ to positions, such that for any reachable term $N$ where $f$ is defined, $f(N)$ is a position of a redex in $N$, and if $f(N)$ is not defined, $N$ is a value.
Using a reduction strategy, a subset of $\Rightarrow$ that isn't nondeterministic, $\Rightarrow_f$, can be defined by $(N,s) \Rightarrow_f (N',s')$ iff $(N,s) \Rightarrow (N',s')$ and $N$ reduces to $N'$ with the redex at $f(N)$.

Essentially what Lemma~\ref{redexDestroyed} demonstrates is that the samples taken during any reduction sequence are independent of each other. This is made more precise in the following lemmas.

@@

\begin{theorem} \label{AstEquivalence}
A closed term $M$ is AST with respect to $\cbv$ iff it is AST.
\end{theorem}
\begin{proof}
For both $\red$ and $\Rightarrow_{\cbv}$, the probability measure on the sample space can be used to define a measure on the reduction sequences of each finite length. The distributions are equal for each length, by induction, as follows.

>>>>

\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{enumitem}
\newcommand{\tY}{\mathsf{Y}}
%\newcommand{\tif}[3]{\textsf{if }#1\textsf{ then }#2\textsf{ else }#3}
\newcommand{\tif}[3]{\mathsf{if}(#1, #2, #3)} % space-saving concrete syntax
\newcommand{\tsample}{\mathsf{sample}}
\newcommand{\tscore}{\mathsf{score}}
\newcommand{\skeletonPlaceholder}{\mathsf{X}} % I'm not sure what the best notation here is.
\DeclareMathOperator{\red}{red}
\DeclareMathOperator{\nnext}{next}
\DeclareMathOperator{\cbv}{cbv}

@@

\subsection{Sampling semantics}
\label{sec:sampling semantics}
This version of the reduction relation allows $\tsample$ to reduce to any number in $[0,1]$. To more precisely specify the probabilities, an additional argument is needed to determine the outcome of random samples. Let $ I = [0,1] \subset \mathbb{R} $, and let $S = I^{\mathbb{N}}$, with the \changed[lo]{Borel $\sigma$-algebra $\calF$} and the probability measure, denoted $\mu$, given by the limit of $1 \gets I \gets I^2 \gets \cdots$, where the maps are the projections that ignore the last element. Equivalently, a basis of measurable sets is $\prod_{i=0}^\infty X_i$ where $X_i$ are all Borel and all but finitely many are $I$, and $\mu (\prod_{i=0}^\infty X_i) = \prod_{i=0}^\infty \Leb(X_i)$.
The maps $\pi_h:S \to I, \; \pi_t:S \to S$ popping the first element are then measurable.
\changed[lo]{Following \cite{DBLP:conf/esop/CulpepperC17}, we call the probability space $(S, \calF, \mu)$ the \emph{entropy space}.}

@@

A \emph{position} 
%\akr{I'm not really sure how to phrase the citation, but this is adapted from \cite{Kennaway96infinitarylambda}.} 
%\lo{The idea is common enough. I think no citation necessary.}
is a finite sequence of steps into a term, defined inductively as
\begin{align*}
P ::= \cdot \mid \lambda ; P \mid @_1 ; P \mid @_2 ; P \mid \underline f_i ; P \mid \tY ; P \mid \textsf{if}_1 ; P \mid \textsf{if}_2 ; P \mid \textsf{if}_3 ; P \mid \tscore ; P.
\end{align*}
The \emph{subterm of $M$ at a position $P$}, denoted $M \mid P$, is defined as
\begin{align*}
M \mid \cdot & = M \\
\lambda x. M \mid \lambda ; P & = M \mid P \\
M_1 M_2 \mid @_i ; P & = M_i \mid P \quad \text{for } i = 1,2 \\
\underline f(M_1,\dots,M_n) \mid \underline f_i ; P & = M_i \mid P \quad \text{for }i \leq n \\
\tY M \mid \tY ; P & = M \mid P \\
\tif{M_1 < 0}{M_2}{M_3} \mid \textsf{if}_i ; P & = M_i \mid P \quad \text{for } i = 1,2,3 \\
\tscore(M) \mid \tscore ; P & = M \mid P
\end{align*}
so that every subterm is located at a unique position, but not every position corresponds to a subterm (e.g. $x \, y \mid \lambda ; \cdot$ is undefined). 
A position such that $M\mid P$ does exist is said to \emph{occur} in $M$. 
%\changed[lo]{
\emph{Substitution} of $N$ at position $P$ in $M$, written $M[N/P]$, is defined similarly.
For example, let 
\[
M = \lambda x \, y. y \, (\tif{x < 0}{y \, (\underline f (x))}{\underline 3})
\qquad
P =\lambda ; \lambda ; @_2 ; \textsf{if}_2 ; @_2
\]
then 
\(
M[\tsample / P] = \lambda x \, y. y \, (\tif{x < 0}{y \, \tsample}{\underline 3}).
\)
%}

Two subterms $N_1$ and $N_2$ of a term $M$, corresponding to positions $P_1$ and $P_2$, can overlap in a few different ways. 
If $P_1$ is an initial segment (i.e.~prefix) of $P_2$ (written as $P_1 \leq P_2$), then $N_2$ is also a subterm of $N_1$. If neither $P_1 \leq P_2$ nor $P_1 \geq P_2$, the positions are said to be disjoint. 
This is mostly relevant in that in any substitution at some position, the subterms at all other positions are unaffected. 
%\lo{The preceding sentence is unclear. Typo? Mostly $\mapsto$ most?}

With this notation, a more general reduction relation $\to$ can be defined. \akr{It would almost work to have one of the neater versions of $\tY$-reduction here.}
%\lo{On first reading, I struggled to parse the following clauses.}
\begin{definition}
\label{def:more general red}
The binary relation $\to$ is defined by the following rules, each is conditional on a redex occurring at position $P$ in the term $M$:
\begin{align*}
  \text{if } M \mid P = (\lambda x.s) v,\ & M \to M[s[v/x]/P] \\
  \text{if } M \mid P = \underline f (\underline r_1, \dots , \underline r_n),\ & M \to M[\underline{f(r_1,\dots,r_n)}/P] \\
  \text{if } M \mid P = \tY \lambda x. s,\ & M \to M[\lambda z. s[(\tY \lambda x. s)/x] z/P] \text{ where $z$ is not free in $s$}\\
  \text{if } M \mid P = \tif{\underline r < 0}{s_1}{s_2},\ & M \to M[s_1/P] \text{ where }r < 0 \\
  \text{if } M \mid P = \tif{\underline r < 0}{s_1}{s_2},\ & M \to M[s_2/P] \text{ where }r \geq 0 \\
  \text{if } M \mid P = \tsample & \text{ and $\lambda$ does not occur after $@_2$ or $\tY$ in $P$},\\ & M \to M[\underline r/P] \text{ where } r \in [0,1] \\
  \text{if } M \mid P = \tscore(\underline r),\ & M \to M[\underline r/P].
\end{align*}
In each of these cases, $M \mid P$ is the \emph{redex}, and reduction takes place at $P$.
\end{definition}

\paragraph{}
Labelling the pre-chosen samples by the positions in the term would also not work because in some cases, a $\tsample$ will be duplicated before being reduced, for example, in $(\lambda x. x \, {\underline 0} \, \mathbin{\underline{+}} \, {x \, \underline 0}) (\lambda y. \tsample)$, both of the $\tsample$ redexes that eventually occur originate at $@_2 ; \lambda$. 
It is therefore necessary to consider possible positions that may occur in other terms reachable from the original term. 
Even this is itself inadequate because some of the positions in different reachable terms need to be considered the same, and the number of reachable terms is in general uncountable, which leads to measure-theoretic issues.

Define a \emph{skeleton} to be a term but, instead of having real constants $\underline r$, it has placeholders $\skeletonPlaceholder$, so that each term has a skeleton and each skeleton can be converted to a term given the appropriate number of real numbers to substitute in. 
\lo{Skeleton is defined in Sec.~\ref{sec:sampling semantics}.}
The positions in a skeleton and the reduction relation on skeletons can be extended from the definitions on terms in the obvious way, with \changed[lo]{$\tif{\skeletonPlaceholder < 0}{A}{B}$} reducing nondeterministically to both $A$ and $B$, \changed[lo]{and $M \to M[\skeletonPlaceholder' / P]$ for a freshly generated placeholder $\skeletonPlaceholder'$, in case $M \mid P = \tsample$ and $\lambda$ does not occur after $@_2$ or $\tY$ in $P$.
(By abuse of notation, we use the same metavariables $M, N$, etc.~for skeletons as for terms.)}
\changed[lo]{For example, we have 
\(
(\lambda x . \tif{x < 0}{x}{\skeletonPlaceholder_1}) \, \tsample
\to
(\lambda x . \tif{x < 0}{x}{\skeletonPlaceholder_1}) \, \skeletonPlaceholder_2
\to
\tif{\skeletonPlaceholder_2 < 0}{\skeletonPlaceholder_2}{\skeletonPlaceholder_1}
\to
\skeletonPlaceholder_i
\) where $i$ is 1 or 2.}

Given a closed term $M$, let $L_0(M)$ be the set of pairs, the first element of which is a \changed[lo]{$\to$-reduction} sequence of skeletons starting at \changed[lo]{the skeleton of} $M$, and the second of which is a position in the final skeleton of the reduction sequence. 
\akr{I've yet to go through and replace references to terms with references to skeletons, but it should be reasonably obvious where the substitution should take place. 
This is potentially even more confusing though because I'm now implicitly identifying terms, skeletons, reduction sequences and skeletal reduction sequences.} 
Reduction sequences are used rather than reachable \changed[lo]{skeletons} because if the same \changed[lo]{skeleton} is reached twice, different samples may be needed, \lo{Give a simple example illustrating this.} but they will be discussed as though they were terms, by abuse of notation.

\lo{Readers would benefit greatly from some examples of an $L_0(M)$-pair, and the relations $\sim_p$ and $\sim_c$, here, before the formal definitions. Explain the underlying intuition.}

\begin{definition}
The relation $\sim$ is defined as the union of the symmetric relations $\sim_p$ \changed[lo]{($p$ for parent-child)} and $\sim_c$ \changed[lo]{($c$ for cousin)} where
\lo{As defined, $\sim_p$ is not a symmetric relation.}
\begin{enumerate}
    \item If $N$ reduces to $O$ with the redex at position $P$ \changed[lo]{(i.e.~$M \to^\ast N \to O = N[\dotDelta / P]$ with $O \mid P = \Delta$ for some redex contraction $\Delta \to \dotDelta$)}, and $Q$ is a position in $N$ disjoint from $P$, then $(N,Q) \sim_p (O,Q)$.

@@

        \item $N \mid P = \tY (\lambda x. A)$, $A$ reduced at $Q$ is $A'$, $(R_i)_i$ is a list of all the positions where $A' \mid R  = x$, ordered from left to right, $O_1$ is $N$ reduced first at $P;\tY;\lambda;Q$ then at $P$, and $O_2$ is $N$ reduced first at $P$ then at $P;\lambda;@_1;Q$ then at $P;\lambda;@_1;R_i;Y;\lambda;Q$ for each $i$ in order.
    \end{enumerate}
    (in which case $O_1$ and $O_2$ are equal as terms, but with different reduction sequences), $O_1'$ and $O_2'$ are the results of applying some reduction sequence to each of $O_1$ and $O_2$ (the same reductions in each case, which is always possible because they're equal terms), and $S$ is a position in $O_1'$ (or equivalently $O_2'$), then $(O_1',S) \sim_c (O_2',S)$.
\end{enumerate}
See \Cref{fig:adk-diagram} for a graphical depiction of $\sim_p$- and $\sim_c$-rules.
\end{definition}

@@

\lo{TODO: Typeset Fig.~\ref{fig:adk-diagram} in vector graphics using PGF/TikZ.}

The reflexive transitive closure $\sim^*$ of this relation is used to define the set of \emph{potential positions} $L(M) = L_0(M) / \sim^*$, and each equivalence class can be considered as the same position as it may occur across multiple reachable \changed[lo]{skeletons}. 
If $(N,P) \sim^* (O,Q)$, then $N \mid P$ and $O \mid Q$ both have the same shape (i.e.~they're either both variables, both applications, both $\tsample$s etc.), therefore it's well-defined to talk of the set of potential positions where there is a $\tsample$, $L_s(M)$. 
\changed[lo]{Formally $L_s(M) := \{[(X, P)]_{\sim^\ast_M} : X \mid P = \tsample\}$.}
The new sample space is then defined as $I^{L_s(M)}$, with the Borel $\sigma$-algebra and product measure.
\changed[lo]{Since $I^{L_s(M)}$ is a countable product, the measure space is well-defined \citep[Cor.~2.7.3]{AshDD00}.}

\iffalse
\changed[lo]{Since $L_s(M)$ is a countable set, there is a unique probability measure $\mu$ on the infinite product $\prod_{i \in L_s(M)} \Sigma_I$ of the Borel $\sigma$-algebras $\Sigma_I$ satisfying
\[
\mu\set{\omega \in \Omega \mid \omega_1 \in A_1, \cdots, \omega_n \in A_n} = \prod_{i=1}^n \Leb(A_i)
\]
where $\Leb$ is the Lebesgue measure on the measurable space $(I, \Sigma_I)$.}
\fi

\paragraph{}
\akr{For now, here's a summary of the lemmas and theorems I intend to prove about this alternative semantics:
\begin{itemize}
    \item If $M \to N$, the relation $\sim_N^*$ defined on $Rch(N)$ is equal to the restriction of $\sim_M^*$ to $Rch(N)$ %\lo{Define $\sim_N^*$.}
    \item Given the definition below of $\Rightarrow$, which is still applicable to this version of $\sim$, and which is the equivalent of $\red$ in this version of the semantics, $\Rightarrow$ is Church-Rosser.
    \item If $M \to N$ at $P$, no position in any term in $Rch(N)$ is related to $(M,P)$ by $\sim_M^*$, and therefore each sample is used at most once in any reduction sequence.
    \item The subprobability distribution of values obtained by reducing a term in the usual call-by-value semantics is equal to that obtained by any reduction strategy on $\Rightarrow$ (i.e.~a sub-relation of $\Rightarrow$ that is a function) (with some allowance made for the fact that some values are not normal forms of $\Rightarrow$, so the reduction strategy might overshoot the value that $\red^\infty$ reaches. I'm not quite sure what form this will take.)
    \item Similarly, the probability of termination is equal in $\red^\infty$ and any reduction strategy on $\Rightarrow$, so almost sure termination can be equivalently defined based on $\Rightarrow$.
    \item For any closed term $N$, if there is any reduction strategy for $N$ and ranking function corresponding to that reduction strategy, $N$ is AST.
    \item Corresponding sparse and antitone-strict ranking function theorems.
\end{itemize}}

@@

\begin{lemma}
\label{lem:sim-M-N}
The relation $\sim$ is defined on $L_0(M)$ with reference to a particular starting term $M$, so different versions, $\sim_M$ and $\sim_N$, can be defined starting at different terms. If $M \to N$, then $\sim^*_N$ is equal to the restriction of $\sim^*_M$ to $L_0(N)$.
\end{lemma}

\begin{proof}
$\sim_N^*$ is trivially a subset of $\sim_M^*$ because $\sim_N$ is a subset of $\sim_M$.

@@

%\begin{definition}
A position in some (reduction sequence of) term(s) is \emph{derived from} another if it is related by the reflexive transitive closure of $\rightsquigarrow$, where $(V,R) \rightsquigarrow (W,T)$ just if $V \to W$ and one of the following cases holds:
\begin{itemize}
    \item $(V,R) \sim_p (W,T)$
    \item $R = T$, $V \to W$ at $A$, and $A > R$
    \item $V | A = (\lambda x.U) Z, U | B = x, R = A;@_2;C$ and $T = A;B;C$
    \item $V | A = \tY (\lambda x.U), R = A;\tY;\lambda;B$ and $T = A;\lambda;@_1;B$
    \item $V | A = \tY (\lambda x.U), U | B = x, R = A;C$ and $T = A;\lambda;@_1;B;C$
\end{itemize}
%\end{definition}

@@

At each reduction step $M \to N$, the sample space must be restricted from $I^{L_s(M)}$ to $I^{L_s(N)}$. 
The injection $L_0(N) \to L_0(M)$ is trivial to define by appending $M \to N$ to each path, and using Lemma~\ref{lem:sim-M-N}, this induces a corresponding injection on the quotient, $L(N) \to L(M)$. 
The corresponding map $L_s(N) \to L_s(M)$ is then denoted $i(M \to N)$.

@@

Let $I$ be $N$ reduced at $P$, and $J$ be $N$ reduced at $P;@_1;\lambda;Q$, so that $N \to I \to O_1$ and $N \to J \to O_2$. All of the reduction positions are $\geq P$, and $G$ is not disjoint from all of them, therefore $G$ is not disjoint from $P$. Let $K$ be the position such that $(N,G) \sim_p (I,K) \sim_p (O_1,H)$. The fact that $(N,G) \sim_p (I,K)$ implies that $G > P;@_1;\lambda$. Let $G = P;@_1;\lambda;L$ and $H = P;L'$. If $L$ is disjoint from $Q$, then $(N,G) \sim_p (J,G) \sim_p (O_2,P;L') = (O_2;H)$ by cases 1 and 2 of $\sim_p$. In the other case, that $L$ isn't disjoint from $Q$, $L > Q$ because none of the positions $<= P;@_1;\lambda;Q$ in $N$ are related to any position in $I$ by $\sim_p$. As $(I,P;L) \sim_p (O_1,P;L')$ (with the redex at $P;Q$), for exactly the same reason $(N,P;@_1;\lambda;L) \sim_p (J,P;@_1;\lambda;L')$ (with the redex at $P;@_1;\lambda;Q$). Because $(N,P;@_1;\lambda;L) \sim_p (J,P;@_1;\lambda;L')$, $N|P;@_1;\lambda;L = J|P;@_1;\lambda;L'$, and $N|P;@_1;\lambda;L \neq $ the variable of $N|P;@_1$, therefore $J|P;@_1;\lambda;L'$ is also not the variable therefore $(J,P;@_1;\lambda;L') \sim_p (O_2, P;L') = (O_2, H)$ by case 2 of $\sim_p$. Combining these results, $(N,G) \sim_p (O_2;H)$ as desired.
\end{proof}

\begin{lemma} \label{lem:redexDestroyed}
If $M \to N$, with the redex at position $P$, then no position in any term reachable from $N$ is related by $\sim^*$ to $(M,P)$.
\end{lemma}
\begin{proof}
Suppose on the contrary that $(M,P) \sim^* (A,D)$, where $N \to^* A$, then by Lemma \ref{simpDoesn'tMix}, $(M,P) \sim_p^* (A,D)$. $M \neq A$ therefore $(M,P) \sim_p$ some position in $N$, but in all the cases of the definition of $\sim_p$, no position in the child term is related to the position of the redex.
\end{proof}

\paragraph{}
The reduction relation $\Rightarrow$ is nondeterministic, so it admits multiple possible reduction strategies. 
A \emph{reduction strategy} starting from a closed term $M$ is a partial function $f$ from $Rch(M)$ to positions, such that for any reachable term $N$ where $f$ is defined, $f(N)$ is a position of a redex in $N$, and if $f(N)$ is not defined, $N$ is a value.
Using a reduction strategy $f$, a subset of $\Rightarrow$ that isn't nondeterministic, $\Rightarrow_f$, can be defined by $(N,s) \Rightarrow_f (N',s')$ just if $(N,s) \Rightarrow (N',s')$ and $N$ reduces to $N'$ with the redex at $f(N)$.

Essentially what Lemma~\ref{lem:redexDestroyed} demonstrates is that the samples taken during any reduction sequence are independent of each other. This is made more precise in the following lemmas.

@@

\begin{theorem} \label{thm:AstEquivalence}
A closed term $M$ is AST with respect to $\cbv$ iff it is AST.
\end{theorem}
\begin{proof}
For both $\red$ and $\Rightarrow_{\cbv}$, the probability measure on the sample space can be used to define a measure on the reduction sequences of each finite length. The distributions are equal for each length, by induction, as follows.
